For submission to the Journal of Integrated Pest Management

### 2021-05-20 editorial_reviews/response_editors_ofv_fife_2021.Rmd

Table 1:
L338: “…with symptoms of orchid fleck virus infection found in…”
I am a little concerned about reporting on plants with symptoms, vs confirmed by diagnostics.  As mentioned earlier, different dichorhaviruses can cause similar symptoms in the same host.  Although remote, there is the possibility another virus could be responsible for the symptoms in a different location

Figure captions:
Please use “orchid fleck virus (OFV)” when describing a physical virus, and "Orchid fleck dichorhavirus" (in italics) when describing the species.

**Response to Reviewer 1:**
**Thank you for reviewing our article. We have made the following changes based on your remarks:**

<!-- - Added discussion on why OFV has not been reported in orchids or citrus if the pathogen and vectors appear so widespread in the state. -->

<!-- - Added discussion of potential control measures to reduce the impact and spread of this virus is warranted. -->

- Removed references to cileviruses

#### Reviewer: 2

**Response to PDF edits (Reviewer #?):**
**Thank you for reviewing our article. We have made the following changes based on your remarks:**

### 2021-05-20 ofv_fife_2021.Rmd
Revised abstract to be less than 250 words and to improve clarity

*Orchid fleck dichorhavirus* (OFV) infects over 50 plant species belonging to the family Orchidaceae, Asparagaceae (Nolinoidaea), and Rutaceae (Citrus). The only known vectors of dichorhaviruses are flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae). OFV was found infecting liriopogons (*Liriope* spp. and *Ophiopogon* spp.),  as well as *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in the landscape in Leon County, Florida, during the summer of 2020. The presence of OFV was confirmed using OFV-specific conventional reverse transcription polymerase chain assay (RT-PCR) assay and Sanger sequencing. RT-PCR amplicons had a 98% identity with known OFV sequences available in NCBI GenBank. Other leaf samples of the plant family Asparagaceae with ringspots and chlorosis were collected from Leon and Alachua counties and tested with PCR and Sanger sequencing to detect OFV. Identification of partial genome sequence from PCR products confirmed the presence of both OFV orchid strains (OFV-Orc1 and OFV-Orc2) in Florida. These strains of OFV are known to infect citrus and cause citrus leprosis disease. Three potential mite vectors were identified from OFV-infected plants using cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (sensu lato), *B. obovatus*, and *B. confusus*. Florida has various native and introduced plants in the landscape that *Brevipalpus* spp. feed on, which are potentially susceptible to OFV. In this study, we report three new hosts from the family Asparagaceae from multiple locations. Our data suggested that *Orchid fleck dichorhavirus* is widely distributed in Florida and might be a threat for various plant species growing in the southeastern US.

@@

\pagebreak
*Orchid fleck dichorhavirus* (OFV) is the type member for the genus *Dichorhavirus*, family *Rhabdoviridae*; a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae) are the only known vectors of dichorhaviruses [@Maeda1998]. *Brevipalpus californicus* (Banks) sensu lato mites are the only group known to transmit OFV in a persistent propagative manner [@Kondo2003].

@@

Subsequent surveys of plants belonging to the subfamily Nolinoidaea in Florida have revealed more sites with Asparagaceae putatively infected with OFV in Leon and Alachua counties (`r citet("tab_1")`). Mites were collected from symptomatic plants in Leon county and observed with phase contrast microscopy. Tenuipalpid mites (flat mites or false spider mites) were commonly found in abundance on the Asparagaceae, which tested positive for OFV. These flat mites were initially identified as *B. californicus* s. l. and later confirmed by the FDACS via Differential Interference Contrast (DIC) microscopy. The *Brevipalpus* mite species complex is known to contain cryptic species [@Childers2011] that require advanced microscopy techniques, such as cryo scanning electron microscopy (Cryo-SEM) for species identification [@Skoracka2015; @Leon2010; @Beard2015]. Additional mite samples were collected from the original OFV detection site, and examined under Cryo-SEM (`r citef("fig_3")`). The determinations approved prior identifications of *B. californicus* s.l. but revealed the presence of two other species *B. obovatus* and *B. confusus* (`r citef("fig_4")`).

@@

### Table 1: List of plants with symptoms of *Orchid fleck dichorhavirus* found in northern Florida

@@

`r fig_nums(name = "fig_4", caption = "Florida is home to other common pest species of *Brevipalpus*, which are potential vectors of *Orchid fleck dichorhavirus*: (a) *B. yothersi*, dorsal (b) *B. yothersi*, lateral (c) *B. obovatus*, dorsal.")`

>>>>

<!-- 250 words or less. -->
<!-- Give scientific name and authority at first mention of each organism. -->
<!-- Do not cite references, figures, tables, probability levels, or results. -->
<!-- Refer to results only in the general sense. -->
<!-- A second abstract in a second language is permitted. -->

We report outbreaks of *Orchid fleck dichorhavirus* (OFV-Orc) infecting three new hosts: *Liriope* spp. *Ophiopogon* and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in Leon and Alachua Counties, Florida. Strains of OFV-Orc infect over 50 plant species belonging to the plant families Orchidaceae, Asparagaceae (Nolinoidaea), and infects *Citrus* (Rutaceae), as citrus leprosis disease. The only known vectors of OFV-Orc are flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae). Florida has various native and introduced plants in the landscape which *Brevipalpus* spp. feed on, which are potentially susceptible to OFV-Orc.

Chlorotic ringspots flecking were seen affecting Liriopogons (*Liriope* spp. and *Ophiopogon* spp.) in Leon County, FL. Nearby *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant viruses, so samples were sent to the Florida Department of Agriculture and Consumer Services for identification. Orchid-infecting *Orchid fleck dichorhavirus* was confirmed via conventional reverse transcription polymerase chain reaction of L-gene (RNA2) and Sanger sequencing of amplicons. Amplicons had a 98% identity with known OFV sequences available in NCBI GenBank.

New samples from symptomatic plants were collected in both Leon and Alachua counties. High-throughput sequencing (HTS) at the USDA-ARS detected OFV-Orc and OFV-Orc2 in *Aspidistra* and *Liriope*, respectively. Three potential mite vectors were identified via cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (Banks) sensu lato, *B. obovatus* Donnadieu, and *B. confusus* Baker.

Our report suggests that *Orchid fleck dichorhavirus* is present in northern Florida, representing a risk for susceptible plants in the southeastern US.

@@

\pagebreak
*Orchid fleck dichorhavirus* (OFV) is the type member for the genus *Dichorhavirus*, family *Rhabdoviridae*; a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae) are the only known vectors of dichorhaviruses [@Maeda1998]. *Brevipalpus californicus* (Banks) sensu lato mites are the only group known to transmit OFV in a persistent propagative manner [@Kondo2003].

@@

Subsequent surveys of plants belonging to the subfamily Nolinoidaea in Florida have revealed more sites with Asparagaceae putatively infected with OFV in Leon and Alachua counties (`r citet("tab_1")`). Mites were collected from symptomatic plants in Leon county and observed with phase contrast microscopy. Tenuipalpid mites (flat mites or false spider mites) were commonly found in abundance on the Asparagaceae, which tested positive for OFV. These flat mites were initially identified as *B. californicus* s. l. and later confirmed by the FDACS via Differential Interference Contrast (DIC) microscopy. The *B. californicus* s. l. species group, sensu @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013], as seen in other *Brevipalpus* species, such as *B. phoenicis* (Geijskes) [@Beard2015].

<!-- You need to identify mites with special techniques [@Skoracka2015] which require supplemental techniques in addition to light microscopy for accurate identification [@Beard2015]. A plea for awareness of cryptic species [@Leon2010] -->

New mite samples were collected from the site where OFV was first confirmed and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Baker and *B. confusus* Donnadieu (`r citef("fig_4")`).

@@

### Table 1: List of Asparagaceae (Nolinoidaea) species with verified cases of *Orchid fleck dichorhavirus*, along with basic descriptions of foliar symptoms seen in the landscape of northern Florida

@@

`r fig_nums(name = "fig_4", caption = "Florida is home to other common pest species of *Brevipalpus*, which are potential vectors of *Orchid fleck dichorhavirus*: (a) *B. phoenicis*, dorsal (b) *B. yothersi*, lateral (c) *B. obovatus*, dorsal.")`
